---
title: "Module 5 Assignment on Resampling"
author: "Jingwen Zhong // Graduate Student"
date: "03/17"
#output: pdf_document
output:
  pdf_document: 
    latex_engine: xelatex
    fig_height: 4
    fig_caption: yes
  df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, error = TRUE, warning = FALSE, message = FALSE, fig.show ='hold',tidy.opts=list(width.cutoff=80))
```

## Module Assignment Questions

## Q1) (*Random Variable Generation*) 

You have learned how to generate exponential random variables using uniform distribution. Now, you will generate normal random variables using one of the methods described in this [LINK](http://interstat.statjournals.net/YEAR/2012/articles/1205003.pdf).

Some useful information from the link: 

- Normal distribution is also called as the Gaussian distribution.For $\mu= 0$ and $\sigma = 1$, we refer to this distribution as the standard normal distribution

- Normal random variables X ’s with mean $\mu$ and variance 2 $\sigma$ are generated by the 
relationship $X = \mu + \sigma Z$ , where Z is the standard normal random variable. 

##### a. As we studied the exponential random variable in the lab session, , generate 1,000 random variables from the standard normal distribution using the method you chose in the link.

```{r}
set.seed(99)
#I use Method 9: Proposed Method
#1. Generate U from the U(0,1) distribution.
unif_rv=runif(n=1000, min=0, max=1)
# Return Z
# Z_from_me = -(log(1/unif_rv-1))/1.702 # log computes natural logarithms(ln=log_e) by default
Z_from_me=0.46615+90.72192*tanh(-31.35694+28.77154*unif_rv)-89.36967*tanh(-2.57136-31.16364*unif_rv)-96.55499*tanh(3.94963-1.66888*unif_rv)+97.36346*tanh(2.31229+1.84289*unif_rv)
```

```{r}
set.seed(99)
#rnorm: The Normal Distribution
Z= rnorm(1000, mean = 0, sd = 1)
```


******
##### b. Check if this generation is verified with the theory: 
1) check if the characteristics (so calculate mean, sd, skewness, kurtosis for each) of empirical and theoretical distributions are similar;
```{r}
hist(Z_from_me,main="Empirical standard normal distributions",prob=TRUE)
curve(dnorm(x,0,1), add=TRUE, col='red',lwd=3)
abline(v = c(mean(Z_from_me),median(Z_from_me)),  col=c("green", "blue"), lty=c(2,2), lwd=c(3, 3))

hist(Z, main="Theoretical standard normal distributions",prob=TRUE)
curve(dnorm(x,0,1), add=TRUE, col='blue',lwd=3)
abline(v = c(mean(Z),median(Z)),  col=c("green", "blue"), lty=c(2,2), lwd=c(3, 3))
```

```{r}

skewness1 = sum((Z_from_me-mean(Z_from_me))^3)/((length(Z_from_me)-1)*sd(Z_from_me)^3)
kurtosis1 = sum((Z_from_me-mean(Z_from_me))^4)/((length(Z_from_me)-1)*sd(Z_from_me)^4)

skewness2 = sum((Z-mean(Z))^3)/((length(Z)-1)*sd(Z)^3)
kurtosis2 = sum((Z-mean(Z))^4)/((length(Z)-1)*sd(Z)^4)

table = matrix(c(mean(Z_from_me), mean(Z), 
                  sd(Z_from_me), sd(Z), 
                  skewness1, skewness2,
                  kurtosis1, kurtosis2),2,4)

colnames(table) = c("mean", "sd", "skewness", "kurtosis")
rownames(table) = c( "Empirical distributions","Theoretical distributions")

knitr::kable(table, caption = " Characteristics check")
```

2) plot a QQ and judge by eye

```{r}
qqplot(Z_from_me, Z)
abline(0,1)
```

3) make a chi-squared test on empirical and theoretical buckets.

```{r}
#use round(min(Z_from_me)) to round(max(Z_from_me)) for normal dist
bins = seq(-4, 4, by=.5)
#df is bins-1: ts=(sum(hist2-hist1)^2/(hist1)), p-value=1-P(X<ts) using chi-sq dist
hist1=hist(Z, freq=TRUE, breaks= bins)$counts + 1
hist2=hist(Z_from_me, freq=TRUE, breaks= bins)$counts + 1
chisq.test(hist2, p=hist1, rescale.p = TRUE)

```

******
##### c. Write an overall comment on your experience.\

I think my empirical distribution is very good. The mean, sd, skewness, kurtosis are all approximate to the theoretical standard normal distributions. Although the p-value of the Chi-sqaure test is too small so that we can reject $H_0: Emperical = Theoretical$ and conclude that my empirical distribution is not standard normal distribution. In fact, I think using the seed(99), my empirical distribution performs better than the theoretical distributions.


\newpage
## Q2) (*Bootstrapping*) 

Let $\hat{\sigma}^2_1$  and $\hat{\sigma}^2_2$ represent the variance estimates of two independent random samples of size 10 and 20, respectively, taken from any normal distributions with variances $\sigma^2_1=12$ and $\sigma^2_2=8$ , respectively. Let a new parameter be defined as $v= \frac{\sigma^2_1}{\sigma^2_2}$. 

Based on randomly generated samples with any choice on means, use bootstrap method (B=1000) to answer each part below: 

```{r}
library(boot)

# define parameters
set.seed(99)
x1 = rnorm(1000,mean=0,sd=sqrt(12))
x2 = rnorm(1000,mean=0,sd=sqrt(8))
df = cbind(x1,x2)
```

##### a. Estimate $\sigma^2_1$ along with the standard error on the estimate. Evaluate if this misses the true value.

```{r}
var1 <- function(data,i){
  var <- sd(data[i])^2 # variance for first normal distribution
  return(var)
}

set.seed(99)
data1 = sample(df[,1], 10)
results <-boot(data=data1, statistic = var1, R=1000)
results

var = results$t
hist(var)

cat("\nit is close to the true value",12)

```

******
##### b. Estimate the parameter $v$ along with the standard error on the estimate. Evaluate if this misses the true value.

```{r}
v= 12/8

var_ratio <- function(data, indices, n1=10, n2=20){
  set.seed(99)
  sample1 = df[sample(size=n1, x=indices, replace = TRUE),1]
  sample2 = df[sample(size=n2, x=indices, replace = TRUE),2]
  ratio = (sd(sample1)/sd(sample2))^2
  return(ratio)}

results2 <- boot(df, statistic = var_ratio, R=1000)
results2

ratio = results2$t[1:1000,]
hist(ratio)

cat("\nit is close to the true value",v)

```

******
##### c. Obtain the percentile-based 95% confidence interval on the $v$ estimate. 

```{r}
boot.ci(results2, type="bca")
```

******
##### d. (BONUS) Search for what theory says about the CI on the $\sigma^2$ and $v$ estimates. Verify if the result in part c convinces the theory.



******
##### e. (BONUS) Can bootstrapping method solve the probability problem, $P(\frac{\hat{\sigma}^2_1}{\hat{\sigma}^2_2}>2)$? Calculate and show. Does it give similar result from the theory?


\newpage
## Q3) (*Advanced Sampling*) 

Watch [`the playlist`](https://youtube.com/playlist?list=PLTpORyMWYJsa-bLjJqzsJKXQCrMAiYu7n) on Hidden Markov, MC, MCMC, Gibbs, Metropolis-Hastings methods (or, use this link <https://youtube.com/playlist?list=PLTpORyMWYJsa-bLjJqzsJKXQCrMAiYu7n>)

##### a. Write one paragraph summary for each MCMC, Gibbs and Metropolis-Hastings method by highlighting how the strategy works (watch six videos, write three paragraphs).

sample from P(x), ( P(x)=f(x)/NC, but we don't know P(x), we only know f(x), NC=normalized constant)

- MCMC: Design a special Markov chain: Learn from the previous samples, and pick the next sample based on that, so samples are no longer independent. Monte carlo: Simulating samples/draws from our target distribution P(x), by simulating this markov chain, and eventually this markov chain is going to simulate draws from P(x) (throw away "burn-in" part).
 
- Gibbs: Use when we sample from a multivariate distribution. Sampling from conditional distributions(not joint distributions), such like P(x|y). For x, y example: Start by initial x, y, then change new x, keep y fixed, then sample new y keep new x fixed, and so on so forth.

- Metropolis-Hastings: step one(candidates): Center a (normal) distribution at the previous sample $x_t$ and with some fixed variance, then sample the next candidate $x_{t+1}$ from that normal distribution. step two: accept the candidate with some probabilities ( $\frac{A(a->b)}{A(b->a)}=\frac{f(b)}{f(a)}\frac{g(a|b)}{g(b|a)}= r_fr_g$, acceptance probablity: $A(a->b)=MIN(1, r_fr_g)$). 


******
##### b. (BONUS) Give an example from `Deep Learning` if any of them is employed.\

HMM is employed in Deep Learning, for example: Siri


\newpage

### Write comments, questions: ...


******
### Disclose the resources or persons if you get any help: ...

### How long did the assignment solutions take?: 10 hrs

## References: ...
